
---
title: "Complete Quarto Tutorial Using RStudio"
author: "Ehsan Zangene"
date: "`r Sys.Date()`"
output: html_document
---

# 1. What is Quarto?
**Quarto** is a next-generation document publishing system (like RMarkdown but more powerful).  
It allows you to create:
- Reports (HTML, PDF, Word)
- Presentations (Reveal.js, Beamer)
- Websites & blogs
- Dashboards & books

Quarto supports **R, Python, Julia, and ObservableJS**.

# 2. Setting up Quarto in RStudio
1. Install **R** and **RStudio**.
2. Install **Quarto CLI** from [https://quarto.org/download/](https://quarto.org/download/).
3. Check installation:
```bash
quarto check
```
4. Install **tinytex** if you want PDF outputs:
```{r}
tinytex::install_tinytex()
```

# 3. Creating Your First Quarto Document
1. In **RStudio**, go to: `File > New File > Quarto Document`
2. Select **HTML**, **PDF**, or **Word** as output.
3. Save it with `.qmd` extension.

# 4. Basic Structure of a Quarto Document
A Quarto file (`.qmd`) has:
1. **YAML Header (Metadata)**:
```yaml
---
title: "My First Quarto Report"
author: "Ehsan Zangene"
date: "2025-07-18"
format: html
---
```
2. **Markdown content**:
```markdown
# Introduction
This is my first Quarto document.
```
3. **Code chunks**:
```{r}
summary(cars)
```

# 5. Running and Rendering
- Click **Render** in RStudio.
- Or use command line:
```bash
quarto render myreport.qmd
```

# 6. Adding R Code
Example:
```{r}
# Simple plot
plot(cars)
```

# 7. Parameters in Quarto
You can define parameters in YAML:
```yaml
params:
  name: "Ehsan"
  n: 10
```
Use them in R:
```{r}
paste("Hello", params$name)
```

# 8. Quarto Presentations
- Use `format: revealjs` for interactive slides:
```yaml
---
title: "Quarto Presentation"
format: revealjs
---
```
- Example slide:
```markdown
## Slide 1
- Point A
- Point B
```

# 9. Quarto + RStudio Workflow
- Integrates like RMarkdown but faster and better cross-language support.
- Supports **interactive HTML widgets** (e.g., `plotly`, `leaflet`).

# 10. Example Document
```markdown
---
title: "Data Analysis with Quarto"
author: "Ehsan"
date: "2025-07-18"
format: html
---

# Introduction
This is a Quarto example.

## Dataset
```{r}
head(mtcars)
```

## Plot
```{r}
library(ggplot2)
ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point()
```
```

# 11. Quarto vs RMarkdown
- Quarto is the future replacement for RMarkdown.
- You don’t need `rmarkdown::render()`, simply `quarto render`.
- Supports more formats like books and interactive notebooks.

# 12. Converting Existing RMarkdown to Quarto
```bash
quarto convert myfile.Rmd
```

# 13. Quarto Websites
- Create a site:
```bash
quarto create-project mysite --type website
cd mysite
quarto render
```
- Add multiple `.qmd` files as site pages.

# 14. Tips & Best Practices
- Use `.qmd` for all reports moving forward.
- Keep R code clean with chunk options:
```{r}
#| echo: false
summary(cars)
```
